---
title: 'Orientation: L1 vs L2'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dev='jpeg',warning=FALSE,message=FALSE, comment="",size="small")
options(formatR.arrow=TRUE, width=100, digits=3,show.signif.stars = TRUE)
library(ggplot2)
library(dplyr)
library(lme4)
#library(lmerTest)
#library(lmtest)
#library(merTools)
library(stringr)
library(latticeExtra)
```

## Data analysis


```{r data, echo=FALSE}
load("orientation.RData")
DLM<-OrientationL1L2

num<-DLM %>% group_by(LangLev,List) %>% summarize(n_distinct(Subject))
```

We have `r sum(num[,3])` subjects in total and `r dim(DLM)[1]`
observations. The subjects are distributed as follows:

```{r echo=FALSE, results=TRUE}
print(num)
```

```{r naomit, echo=FALSE}
DLM<-na.omit(DLM)
DLM$Subject<-factor(DLM$Subject)
num<-DLM %>% group_by(LangLev,List) %>% summarize(n_distinct(Subject))
```

After removal of entries with missing data due to wrong answers, we have
`r sum(num[,3])` subjects in total and `r dim(DLM)[1]` observations
left. The subjects are distributed as follows:


```{r echo=FALSE, results=TRUE}
print(num)
```

```{r dataoverview, eval=FALSE, echo=FALSE}
# Let's have a glimpse at the data:
# tbl_df(DLM)
glimpse(DLM)
```

## Model Building

We set up a categorical variable, `LangLev`, to inform about the
speakers L1 and L2 and a another categorical variable, `Type`, to inform
about the `Type` of question MATCH and MISMATCH. Further, each different
question is identified by the variable `Trial` that we treat also as a
categorical variable, since we assume no influence of the order in which
the question was posed. We record in the variable `List` the list from
which the question is drawn, that is, list A or list B.  The response
variable is reaction time measured in milliseconds.

We are not intersted in the effect of the specific trials or the
specific subjects. However, we are interested in knowing whether these
effects are important for the variance of responses. Hence, we specify a
mixed effects model with *random effects* associated to the factors
`Trial` and `Subject` and *fixed effects* associated with the other
factors. Due to the fact that we split trials between two lists and
administered the two lists to different subjects the random effects of
`Trial` and `Subject` are nested within `List`. Further, for the way we
split trials, `Trial` is nested also within `Type`. Within the nesting
structure subjects and questions are fully crossed. We fit our model in
R using the package `lme4`
[Douglas Bates, Martin Maechler, Ben Bolker, Steve Walker (2015). Fitting Linear Mixed-Effects Models Using lme4. Journal of Statistical Software, 67(1), 1-48. <doi:10.18637/jss.v067.i01>].
As far as the fixed effects are concerned, we are interested in the
fixed effects of `LangLev`, `Type` and their interaction. We expect no
effect by `List` but we include it in the model to test our expectation.


The distribution of reaction times has longer tail on the right and never reaches
zero. Under these conditions, logarithm transformation or generalized
linear mixed models with Gamma family and log link function are common
approaches to meet the assumptions underlying mixed linear models. In
our case, we found that a log-transformed response variable improved
considerably the diagnostic plots (conditional and marginal residual
qqplots) of the selected models. A Gamma family transformation with log
link function yield much worse diagnostic plots.


Hence, considering the design explained above and the goals of this
study we analyze the following model is: 

```{r basicmodel, warning=TRUE}
lmm<- lmer(log(Time) ~ Type + LangLev + List + Type:LangLev + (1 | List:Subject)+ (1 | List:Type:Trial),
           data = DLM, REML=FALSE)
```

## Likelihood Ratio Tests

The random effects are significant. 

```{r randomeffectslrt, warning=FALSE, eval=TRUE, echo=TRUE}
lm.0 <- lm(log(Time) ~ 1, data = DLM)
# LeaYrs AECI AMGE
lmm.0 <- lmer(log(Time) ~ (1 | List:Subject)+ (1 | List:Subject:Trial), data = DLM)
AIC(lm.0,lmm.0)

lmm.4 <- update(lmm,.~.-(1 | List:Subject))
anova(lmm.4,lmm)

lmm.5 <- update(lmm,.~.-(1 | List:Type:Trial))
anova(lmm.5,lmm)
```

As far as fixed effects are concerned, only `LangLev` is significant:

```{r mixedmodels2, warning=FALSE, eval=TRUE, echo=TRUE}
#lmm.0 <- lmer(log(Time) ~ (1 | List:Subject)+ (1 | List:Type:Trial), data = DLM)
lmm.1 <- update(lmm, .~. - Type - Type:LangLev)
anova(lmm,lmm.1)

lmm.2 <- update(lmm, .~. - LangLev - Type:LangLev)
anova(lmm,lmm.2)

lmm.3 <- update(lmm, .~. - Type:LangLev)
anova(lmm,lmm.3)

lmm.4 <- update(lmm, .~. - List)
anova(lmm,lmm.4)
```

## Analysis of Variance


In the following, we report the Anova of type III with Satterwhite
approximation to take into account the different sizes of samples in the
experimental units. We note that in the literature the use of this
procedure is controversial, however, we observe here that the results
confirm those of likelihood ratio tests.


Note that due to the removal of entries with missing values, our
experimental set up is not balanced. Hence, it is controversial to
report p-values for F or t statistics as there are no analytical results
for null distributions of parameter estimates in complex situations
(e.g., unbalanced or partially crossed designs) [Bates 2015]. 

The package `lmerTest` makes availbale an analysis of variance of Type
III with Satterthwaite approximation for degrees of freedom.


```{r anova, message=FALSE, warning=FALSE}
require(lmerTest)
lmm<- lmer(log(Time) ~ Type + LangLev + List + Type:LangLev + (1 | List:Subject)+ (1 | List:Type:Trial),
           data = DLM, REML=FALSE)
anova(lmm, type=3)
```
```{r results='asis', echo=FALSE}
#library(xtable)
#print(xtable(anova(lmm)),type="html")
```

and the effects with the t test:

```{r sum}
summary(lmm, transform="log")
```


We observe that the two random effects, subjects and trials, account for
a good portion of the standard deviation of the residuals 
`r attr(VarCorr(lmm),"sc")`, 
namely, for
`r attr(VarCorr(lmm)$"List:Subject","stddev")` and 
`r attr(VarCorr(lmm)$"List:Type:Trial","stddev")`, 
respectively. The model
is translated by means of a significant intercept. The estimated effects
are in the log transformed scale.


These results on the statistical significance of the terms are confirmed
also by backward elimination of non-significant effects and by boostrap
confidence intervals:
```{r confintstd, echo=TRUE, eval=TRUE}
#confint(lmm,method="Wald")
confint(lmm,method="boot")
```

We also observed the analysis by the `merTools` package that calculates
simulated values of the (fixed and random) effects from the posterior
using the function `arm::sim`:
```{r simulconfint, height=6, width=4, echo=TRUE}
detach("package:lmerTest",unload=TRUE)
require(merTools)
lmm <- lmer(formula = log(Time) ~ Type + LangLev + List + Type:LangLev + (1 | List:Subject) + 
    (1 | List:Type:Trial), data = DLM, REML = FALSE)
plotREsim(REsim(lmm, n.sims= 100), level=0.95, stat="mean", sd = TRUE)
plotFEsim(FEsim(lmm, n.sims = 100), level = 0.95, stat = 'mean', intercept = FALSE, sd=TRUE)
```

These plots also confirm our conclusions that `Type`, `List` and
`Type:LangLev` are not significant.


## Conclusions

We plot the random and the fixed effects together with their 95%
confidence levels in the figure. The confidence levels are obtained by
the estimated standard deviation and the normal distribution. However,
other analysis with boostrapped confidence intervals confirmed the same
conclusions.


```{r randomeff, height=6, width=4, echo=FALSE}
#plotREsim(REsim(lmm, n.sims = 100), stat = 'median', sd = TRUE)
source("lib.R")
#randeff.plot(lmm,"Subject")
#randeff.plot(lmm,"Type:List:Type:Trial")
randeff.plot2(lmm)
```


```{r fixedplot, height=4, width=5, echo=FALSE}
require("effects")

plot(effects::Effect(c("LangLev"), lmm, transformation=list(link=log, inverse=exp)),cex=1,width=0,lwd=1,ylab="Response time [ms]",xlab="Language level",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),colors=c("grey35","black"))

plot(effects::Effect(c("Type"), lmm, transformation=list(link=log, inverse=exp)),cex=1,width=0,lwd=1,ylab="Response time [ms]",xlab="Type",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),colors=c("grey35","black"))

plot(effects::Effect(c("List"), lmm, transformation=list(link=log, inverse=exp)),cex=1,width=0,lwd=1,ylab="Response time [ms]",xlab="List",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),colors=c("grey35","black"))

plot(effects::Effect(c("LangLev", "Type"), lmm, transformation=list(link=log, inverse=exp)),x.var="Type",cex=1,width=0,lwd=1,ylab="Response time [ms]",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),as.table=TRUE,colors=c("grey35","black")) 
```


From the figure we can observe that the random effects cause often a
significant effect by shifting the intercept by a quantity larger than
zero. For the fixed effects we transformed back the effects in the
linear time scale of milliseconds. It is evident that L1 speakers have a
shorter reaction time and that `Type` and `List` do not have a
significant impact.


<!--
## op <- options(contrasts = c("contr.sum", "contr.poly"))
# Type + CoR + Hand + EO + List + CEF + SRRC + Pre + Post + Post2 + STAY + LeaYrs + DUH + RPV + AMGE + AECI, 
-->



## Diagnostic plots

```{r plotlmr, fig.height=8, echo=FALSE}
#lmm<- glmer(Time ~ LangLev + Type + (1 | Subject)+ (1 | Type:List:Type:Trial), data = DLM,family=Gamma(link = "log"))
par(mfrow=c(2,2))
# plot(lm4,which=1:4)

plot(fitted(lmm, Type = "response"), residuals(lmm, Type = "response"),
     main = "Conditional residuals", xlab = "Predicted", ylab = "Residuals")

res <- residuals(lmm, Type = "response")
qqnorm(res, main = "Conditional residuals, QQplot")
qqline(res)

lm.0 <- lm(log(Time) ~  LangLev + Type + List + Type:LangLev, data = DLM)
x <- model.matrix(lm.0)
pred <- x %*% fixef(lmm)
res <- DLM$Time - pred
plot(pred, res, main = "Marginal residuals", xlab = "Predicted", ylab = "Residuals")
qqnorm(res, main = "Marginal residuals, QQplot")
qqline(res)

```

```{r plot1, fig.width=3, echo=FALSE, eval=FALSE}
plot(lmm,Type=c("p","smooth"))
plot(lmm,sqrt(abs(resid(.))) ~ fitted(.), Type=c("p","smooth"))
qqmath(lmm,id=0.005)
# package HLMdiag influence.ME
```


The joint qqplot looks normal. The marginal look less compliant with
the normality assumptions. 


