---
title: 'Size: L1 vs L2'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dev='jpeg',warning=FALSE,message=FALSE, comment="",size="small")
options(formatR.arrow=TRUE, width=100, digits=3,show.signif.stars = TRUE)
library(ggplot2)
#library(party)
library(dplyr)
library(lme4)
#library(lmerTest)
#library(lmtest)
library(stringr)
#library(merTools)
library(latticeExtra)
```

## Data Analysis


```{r data, echo=FALSE}
load("size.RData")
DLM<-SizeL1L2

num<-DLM %>% group_by(LangLev,List) %>% summarize(n_distinct(Subject))
```

We have `r as.numeric(num[2,2])` subjects for L1 and `r
as.numeric(num[1,2])` for L2 for a total of `r dim(DLM)[1]`
observations. The subjects are distributed as follows:

```{r echo=FALSE, results=TRUE}
print(num)
```

```{r naomit, echo=FALSE}
DLM<-na.omit(DLM)
DLM$Subject<-factor(DLM$Subject)
num<-DLM %>% group_by(LangLev,List) %>% summarize(n_distinct(Subject))
```

After removal of entries with missing data due to wrong answers, 
we have `r sum(num[,3])` subjects in total
and `r dim(DLM)[1]` observations left. The
subjects are distributed as follows:

```{r echo=FALSE, results=TRUE}
print(num)
```


```{r dataoverview, eval=FALSE, echo=FALSE}
# Let's have a glimpse at the data:
# tbl_df(DLM)
glimpse(DLM)
```

## Model Building

The set up is the same as for the Orientation L1 vs L2 case. See there for
details.

Hence, we consider the following model:

```{r basicmodel, warning=TRUE}
lmm<- lmer(log(Time) ~ Type + LangLev + List + LangLev:Type + (1 | List:Subject)+ (1 | List:Type:Trial), data = DLM, REML=FALSE)
```



### Likelihood Ratio Tests

For the random factors:

```{r randomeffectslrt, warning=FALSE, eval=TRUE, echo=TRUE}
lm.0 <- lm(log(Time) ~ 1, data = DLM)
# LeaYrs AECI AMGE
lmm.0 <- lmer(log(Time) ~ (1 | List:Subject)+ (1 | List:Subject:Trial), data = DLM)
AIC(lm.0,lmm.0)

lmm.4 <- update(lmm,.~.-(1 | List:Subject))
anova(lmm.4,lmm)

lmm.5 <- update(lmm,.~.-(1 | List:Type:Trial))
anova(lmm.5,lmm)
```

and for the fixed factors: 

```{r mixedmodels2, warning=FALSE}
lmm.0 <- lmer(log(Time) ~ (1 | List:Subject)+ (1 | List:Type:Trial), data = DLM)
lmm.1 <- update(lmm.0, .~. + List)
anova(lmm.0,lmm.1)

lmm.2 <- update(lmm.0, .~. + Type)
anova(lmm.0,lmm.2)

lmm.3 <- update(lmm.0, .~. + LangLev)
anova(lmm.0,lmm.3)

lmm.4 <- update(lmm.0, .~. + List + Type + LangLev)
lmm.5 <- update(lmm.4, .~. + LangLev:Type)
anova(lmm.4,lmm.5)
```

We see that `Type` and `LangLev` are significant. 




## Analysis of Variance

We present here the Anova of type III with Satterwhite
approximation. Again we refer to the Orientation L1 vs L2 case for more
details.

```{r anova, message=FALSE, warning=FALSE}
require(lmerTest)
lmm<- lmer(log(Time) ~ Type + LangLev + List + LangLev:Type + (1 | List:Subject)+ (1 | List:Type:Trial), data = DLM, REML=FALSE)
anova(lmm)
```

and the effects with t test:
```{r sum}
summary(lmm, transform="log")
```


We seek confirmation of results by
Bootstrap confidence intervals

```{r confintstd, eval=TRUE, echo=TRUE}
#confint(lmm,method="Wald")
confint(lmm,method="boot")
```

We also include the analysis via `merTools` package that calculates
simulated values of the (fixed and random) effects from the posterior
using the function `arm::sim`:
```{r simulconfint, height=6, width=4, echo=TRUE}
detach("package:lmerTest",unload=TRUE)
require(merTools)
lmm <- lmer(formula = log(Time) ~ LangLev + Type + List + LangLev:Type + (1 | List:Subject) + 
     (1 | List:Type:Trial), data = DLM, REML = FALSE)
plotREsim(REsim(lmm, n.sims= 100), level=0.95, stat="mean")
plotFEsim(FEsim(lmm, n.sims = 100), level = 0.95, stat = 'mean', intercept = FALSE)
```

These plots also confirm our conclusions that:


## Conclusions



We plot the random and the fixed effects together with their 95%
confidence levels in the figure. The confidence levels are obtained by
the estimated standard deviation and the normal distribution. However,
other analysis with boostrapped confidence intervals confirmed the same
conclusions. For the fixed effects we tranform back the measurments in
the original linear scale of milliseconds.

```{r randomeff, height=6, width=4, echo=FALSE}
#plotREsim(REsim(lmm, n.sims = 100), stat = 'median', sd = TRUE)
source("lib.R")
#randeff.plot(lmm,"Subject")
#randeff.plot(lmm,"Type:List:Type:Trial")
randeff.plot2(lmm)
```



```{r fixedplot, height=4, width=5, echo=FALSE}

plot(effects::Effect(c("LangLev"), lmm, transformation=list(link=log, inverse=exp)),cex=1,width=0,lwd=1,ylab="Response time [ms]",xlab="Language level",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),colors=c("grey35","black"))

plot(effects::Effect(c("Type"), lmm, transformation=list(link=log, inverse=exp)),cex=1,width=0,lwd=1,ylab="Response time [ms]",xlab="Type",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),colors=c("grey35","black"))

plot(effects::Effect(c("List"), lmm, transformation=list(link=log, inverse=exp)),cex=1,width=0,lwd=1,ylab="Response time [ms]",xlab="List",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),colors=c("grey35","black")) 


plot(effects::Effect(c("LangLev", "Type"), lmm, transformation=list(link=log, inverse=exp)),x.var="Type",cex=1,width=0,lwd=1,ylab="milliseconds",main="", bg="grey50",fg="black",alternating=FALSE,par.settings = ggplot2like(),lines.title=0,between=list(x=1),as.table=TRUE,colors=c("grey35","black")) 
```



We conclude that L1 speakers have shorter reaction times than L2
speakers under both experimental conditions of MATCH and MISMATCH
questions. In addtion, MISMATCH questions seem slightly more difficult
to be answered. This effect was shown to be significant from a bootstrap
analysis on the transformed data.


## Diagnostic plots

```{r plotlmr, fig.height=8}
#lmm<- glmer(Time ~ LangLev + Type + (1 | Subject)+ (1 | Type:List:Type:Trial), data = DLM,family=Gamma(link = "identity"))
par(mfrow=c(2,2))
# plot(lm4,which=1:4)

plot(fitted(lmm, Type = "response"), residuals(lmm, Type = "response"),
     main = "Conditional residuals", xlab = "Predicted", ylab = "Residuals")

res <- residuals(lmm, Type = "response")
qqnorm(res, main = "Conditional residuals, QQplot")
qqline(res)

lm.0 <- lm(log(Time) ~ Type + LangLev + List + LangLev:Type, data = DLM)
x <- model.matrix(lm.0)
pred <- x %*% fixef(lmm)
res <- DLM$Time - pred
plot(pred, res, main = "Marginal residuals", xlab = "Predicted", ylab = "Residuals")
qqnorm(res, main = "Marginal residuals, QQplot")
qqline(res)

```

```{r plot1, fig.width=3, echo=FALSE, eval=FALSE}
plot(lmm,Type=c("p","smooth"))
plot(lmm,sqrt(abs(resid(.))) ~ fitted(.), Type=c("p","smooth"))
qqmath(lmm,id=0.005)
# package HLMdiag influence.ME
```


The joint qqplot looks normal. The marginal look less compliant with the
normality assumptions. 

